The Changing Shape of Empire: A GM Study of Chimú Bottles in Museum Collections

### 2023-01-05 supp/01-linear.Rmd
<edit index>

```{r load packages 3, echo=TRUE}
# install required analysis packages
#devtools::install_github("tidyverse/tidyverse")
#devtools::install_github("mlcollyer/RRPP")
#devtools::install_github("kassambara/ggpubr")
#devtools::install_github('sinhrks/ggfortify')
# load libraries
library(here)
library(tidyverse)
library(RRPP)
library(ggpubr)
library(ggfortify)
library(cluster)
library(wesanderson)
library(ggExtra)
```

@@

## Analyses of Variance (ANOVA) for _variable_ ~ _group_

### ANOVA _Orifice diameter_ ~ _group_

```{r anova.gp.od, out.width = "100%", dpi = 300, echo=TRUE}
# anova = orifice diameter ~ group
gp.od <- lm.rrpp(od ~ gp, 
                 SS.type = "I", 
                 DATA = DATA, 
                 iter = 9999, 
                 print.progress = FALSE)
anova(gp.od)
```

### ANOVA _Base diameter_ of Chimú vessels

```{r anova.gp.bd, out.width = "100%", dpi = 300, echo=TRUE}
# anova = base diameter ~ group
gp.bd <- lm.rrpp(bd ~ gp, 
                 SS.type = "I", 
                 DATA = DATA, 
                 iter = 9999, 
                 print.progress = FALSE)
anova(gp.bd)
```

### ANOVA _Neck diameter_ of Chimú vessels

```{r anova.gp.nd, out.width = "100%", dpi = 300, echo=TRUE}
# anova = neck diameter ~ group
gp.nd <- lm.rrpp(nd ~ gp, 
                 SS.type = "I", 
                 DATA = DATA, 
                 iter = 9999, 
                 print.progress = FALSE)
anova(gp.nd)
```

### ANOVA _Belly diameter_ of Chimú vessels

```{r anova.gp.bl, out.width = "100%", dpi = 300, echo=TRUE}
# anova = belly diameter ~ group
gp.bl <- lm.rrpp(bl ~ gp, 
                 SS.type = "I", 
                 DATA = DATA, 
                 iter = 9999, 
                 print.progress = FALSE)
anova(gp.bl)
```

### ANOVA _Height_ of Chimú vessels

```{r anova.gp.h, out.width = "100%", dpi = 300, echo=TRUE}
# anova = height ~ group
gp.h <- lm.rrpp(h ~ gp, 
                SS.type = "I", 
                DATA = DATA, 
                iter = 9999, 
                print.progress = FALSE)
anova(gp.h)
```

### ANOVA _Rim Thickness_ of Chimú vessels

```{r anova.gp.rt, out.width = "100%", dpi = 300, echo=TRUE}
# anova = rim thickness ~ group
gp.rt <- lm.rrpp(RT ~ gp, 
                 SS.type = "I", 
                 DATA = DATA, 
                 iter = 9999, 
                 print.progress = FALSE)
anova(gp.rt)
```

>>>>

```{r load packages 3, echo=TRUE}
# load libraries
library(here)
library(tidyverse)
library(vegan)
library(RVAideMemoire)
library(ggpubr)
library(ggfortify)
library(cluster)
library(wesanderson)
library(ggExtra)
```

@@

## MANOVA

```{r manova}
# perMANOVA
ceramic.matrix <- as.matrix(DATA[,3:8]) # response vars
ceramic.mat <- sqrt(ceramic.matrix) # sqrt transform
ceramic.dist <- vegdist(ceramic.mat,
                        method = 'bray') # dissimilarity matrix
set.seed(5) # make results reproducible

# model: morphology as a function of form (jar v. bottle)
ceramic.manova <- adonis2(ceramic.dist ~ form,
                          data = DATA,
                          permutations = 10000,
                          method = 'bray')

## do bottles and jars differ?
ceramic.manova
```
